---
title: 'Week 30: Animal Outcomes'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r load libraries, include=FALSE}
library(tidyverse)
library(tidytuesdayR)
library(cowplot)
library(ggalluvial)
library(extrafont)

loadfonts(device = "win", quiet = TRUE)
theme_set(theme_cowplot())
```

```{r load data, include=FALSE}
tuesdata <- tidytuesdayR::tt_load(2020, week = 30)
tuesdata

animal_outcomes <- tuesdata$animal_outcomes
```

## Data Exploration

Simple counting to see how the data looks and feels

```{r}
animal_outcomes %>% 
  count(year, wt = Total) %>% 
  ggplot(aes(year, n)) +
  geom_line()

animal_outcomes %>% 
  count(animal_type, wt = Total)
```

## Pivot-longer

Make the dataset longer and get it visualization ready.

Lumped the following categorical variables:

* **animal_type** as Cats, Dogs and Other
* **Outcome** as Euthanized, Reclaimed, Rehomed and Other

```{r}
animal_outcomes_longer <- animal_outcomes %>% 
  pivot_longer(cols = ACT:WA, names_to = "States") %>% 
  select(-Total) %>% 
  mutate(value = replace_na(value, 0),
         animal_type = fct_lump(animal_type, n = 2, w = value),
         animal_type = fct_reorder(animal_type, value),
         outcome = fct_lump(outcome, 3, w = value))
```

Some initial conclusions:

At an aggregate level, over the years:

* More cats than dogs were euthanized and rehomed.
* More dogs than cats were reclaimed

```{r}
bar_plot <- animal_outcomes_longer %>% 
  filter(outcome != "Other") %>% 
  count(animal_type, outcome, wt = value) %>% 
  ggplot(aes(animal_type, n, fill = outcome)) +
  geom_col(alpha = 0.8) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::comma_format(scale = 1/1000, suffix = "k")) + 
  coord_flip() +
  theme_minimal_vgrid(12) +
  ggthemes::scale_color_tableau() +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.justification = "center",
        axis.title = element_blank())
```

The above plot is based on the aggregate data over all years. The next graph visualizes this year-wise to decipher any trend. 

Also, it will be better to look at proportions rather than absolute numbers.

```{r}
line_plot <- animal_outcomes_longer %>% 
  group_by(year, animal_type) %>% 
  mutate(total = sum(value)) %>% 
  ungroup() %>% 
  group_by(year, animal_type, outcome) %>% 
  mutate(value = sum(value)) %>%
  summarise(prop = first(value / total), .groups = "drop") %>% 
  filter(outcome %in% c("Euthanized", "Reclaimed", "Rehomed"),
         animal_type %in% c("Dogs", "Cats")) %>% 
  ggplot(aes(year, prop, color = animal_type)) +
  geom_line() +
  facet_wrap(~outcome) +
  scale_x_continuous(breaks = c(1999, 2007, 2015)) +
  scale_y_continuous(labels = scales::percent_format()) +
  panel_border() +
  ggthemes::scale_color_tableau() +
  theme_minimal_grid(10) +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        legend.justification = "center",
        strip.text.x = element_text(size = 12, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10, face = "bold"),
        panel.spacing.x = unit(1, "lines"),
        panel.grid.major.y = element_line(color = "grey85", size = 0.5)) +
  labs(y = "Proportion")
```

```{r}
alluvial_plot <- animal_outcomes_longer %>% 
  group_by(year, animal_type) %>% 
  mutate(total = sum(value)) %>% 
  ungroup() %>% 
  group_by(animal_type, outcome) %>% 
  summarise(value = sum(value), .groups = "drop") %>% 
  filter(outcome %in% c("Euthanized", "Reclaimed", "Rehomed"),
         animal_type %in% c("Dogs", "Cats")) %>% 
  mutate(across(where(is.factor), as.character)) %>% 
  ggplot(aes(y = value, axis1 = animal_type, axis2 = outcome)) +
  geom_alluvium(aes(fill = outcome)) +
  geom_stratum(size = 0.75, color = "white", aes(fill = outcome)) +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            size = 5, color = "white") +
  theme_nothing() 
```


```{r}
title <- ggdraw() +
  draw_label("Dogs reclaimed often, Cats euthanized more",
             fontface = "bold",
             fontfamily = "Stencil",
             size = 18,
             x = 0.3,
             hjust = 0)
first_col <- plot_grid(bar_plot, line_plot, nrow = 2)
two_plots <- plot_grid(first_col, alluvial_plot, ncol = 2, 
          rel_widths = c(1.75, 1))
final_plot <- plot_grid(title, two_plots, ncol = 1,
          rel_heights = c(0.1, 2))

final_plot %>% 
  theme(back)
```

# Save final plot

```{r}
ggsave("2020/plots/week_30.png", final_plot, device = "png", width = 15, height = 7)
```


```{r include=FALSE}
knitr::knit_exit()
```

## Appendix

Quick sum to check that Total matches the individual entries

```{r}
animal_outcomes_test <- animal_outcomes %>% 
  rowwise() %>% 
  mutate(total_calc = sum(across(.cols = ACT:WA), na.rm = TRUE),
         check = Total - total_calc)
```

```{r}
animal_outcomes %>% 
  summarise(across(.fns = ~ sum(is.na(.))))
```

```{r}
temp <- animal_outcomes %>% 
  rowwise() %>% 
  filter(across(.fns = ~ all(!is.na(.))))

animal_outcomes_test %>% 
  anti_join(temp)
```

```{r}
animal_outcomes %>% 
  rowwise() %>% 
  mutate(total_calc = sum(across(.cols = ACT:WA), na.rm = TRUE),
         check = Total - total_calc) %>% 
  filter(year == 2005, animal_type == "Dogs", outcome == "In Stock")
```

```{r}
animal_outcomes <- tuesdata$animal_outcomes

animal_outcomes_test <- animal_outcomes %>% 
  rowwise() %>% 
  mutate(total_calc = sum(across(.cols = ACT:WA), na.rm = TRUE),
         check = Total - total_calc)

animal_outcomes_test %>% 
  anti_join(temp)
```

```{r}
animal_outcomes %>% 
  gather(key = state, value = animals, ACT:WA) %>% 
  # filter(is.na(animals))
  mutate(animals = replace_na(animals, 0))
```

```{r}
df <- animal_outcomes %>% 
  gather(key = state, value = animals, ACT:WA) %>% 
  mutate(animals = replace_na(animals, 0),
         animal_type_fct = fct_lump(animal_type, 3, w = animals))

df %>% 
  count(animal_type_fct, wt = animals) %>% 
  ggplot(aes(animal_type_fct, n)) +
  geom_col() +
  coord_flip()
```

## Time-series plots

```{r}
df <- animal_outcomes %>% 
  gather(key = state, value = animals, ACT:WA) %>% 
  mutate(animals = replace_na(animals, 0),
         animal_type_fct = fct_lump(animal_type, 3, w = animals))

df %>% 
  filter(outcome == "Reclaimed") %>% install.packages("ggalluvial")
  ggplot(aes(x = year, y = animals)) +
  geom_path(aes(color = animal_type_fct)) +
  facet_wrap(~state)
```

```{r}
animal_outcomes %>%
  filter(animal_type %in% c("Cats", "Dogs")) ->
  cats_dogs

ggplot(cats_dogs, aes(x= year, y = animals, col = outcome)) +
  geom_line() + 
  facet_grid(state ~ animal_type)+
  theme_minimal()
```


## Alluvial diagrams

```{r}
# install.packages("ggalluvial")
library(ggalluvial)

df <- animal_outcomes %>% 
  gather(key = state, value = animals, ACT:WA) %>% 
  mutate(animals = replace_na(animals, 0),
         animal_type_fct = fct_lump(animal_type, 3, w = animals))

plot <- df %>% 
  group_by(animal_type_fct, outcome) %>% 
  summarise(total = sum(animals)) %>% 
  ggplot(aes(axis1 = animal_type_fct, axis2 = outcome,
             y = total)) +
  geom_alluvium(aes(fill = outcome)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()

ggsave("2020/plots/week_30.png", plot, device = "png", width = 15, height = 8)
```

### Testing

```{r}
titanic_wide <- data.frame(Titanic)
head(titanic_wide)

ggplot(data = titanic_wide,
       aes(axis1 = Class, axis2 = Sex, axis3 = Age,
           y = Freq)) +
  scale_x_discrete(limits = c("Class", "Sex", "Age"), expand = c(.2, .05)) +
  xlab("Demographic") +
  geom_alluvium(aes(fill = Survived)) +
  geom_stratum() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal()
```

