---
title: 'Week 29: Astronauts'
output: html_document
---

## Load required packages

```{r lib}
library(tidyverse)
library(waffle)

theme_set(theme_light())
```

## Download and import the raw data

```{r data}
tuesdata <- tidytuesdayR::tt_load(2020, week = 29)
tuesdata
astronauts_data <- tuesdata$astronauts
```

## Helper functions

```{r utils}
count_prop <- function(df, ...) {
  count(df, ...) %>% 
    mutate(prop = n / sum(n))
}
```

## Data exploration

### Distribution by sex

About 90% of the data is related to male astronauts.

However, interestingly amongst female astronauts more than 70% are civilians while amongst males the proportion of civilians is just 35%.

```{r explore}
astronauts_data %>% 
  count_prop(sex)

astronauts_data %>% 
  group_by(sex) %>% 
  count_prop(military_civilian)
```

### Which countries do the astronauts represent?

```{r}
astronauts_data %>% 
  count_prop(nationality, sort = TRUE) %>% 
  top_n(5) %>% 
  mutate(nationality = fct_reorder(nationality, prop)) %>% 
  ggplot(aes(nationality, prop)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format()) +
  expand_limits(y = c(0, 1)) +
  coord_flip()
```

### What is the avergae waiting time between selection and first mission?

```{r}
first_mission <- astronauts_data %>% 
  filter(mission_number == 1) %>% 
  mutate(years_waiting = year_of_mission - year_of_selection) %>% 
  filter(years_waiting > 0) 

first_mission %>% 
  ggplot(aes(years_waiting)) +
  geom_density(fill = "black", alpha = 1/3) +
  expand_limits(x = 0, y = 0)
```

## Space mission over the years US v/s Russia

```{r}
missions_by_year <- astronauts_data %>% 
  mutate(nationality = fct_lump(nationality, 2)) %>% 
  group_by(year_of_mission, nationality) %>% 
  summarise(num_missions = length(unique(mission_title))) 

missions_by_year %>%  
  filter(nationality != "Other") %>% 
  ggplot(aes(year_of_mission, num_missions)) +
  geom_line(aes(group = nationality, color = nationality)) +
  scale_x_continuous(breaks = seq(1960, 2020, 5)) +
  expand_limits(y = 0)
```

## Gap between mission for astronauts with multiple space visits

```{r}
multiple_missions <- astronauts_data %>% 
  mutate(nationality = fct_lump(nationality, 2)) %>% 
  filter(total_number_of_missions > 1) %>% 
  group_by(name) %>% 
  arrange(mission_number) %>% 
  mutate(on_earth = year_of_mission - lag(year_of_mission, default = min(year_of_mission))) %>%
  select(nationality, year_of_mission, mission_number, total_number_of_missions, on_earth) %>% 
  arrange(name) %>% 
  ungroup()

names <- multiple_missions %>% 
  arrange(desc(total_number_of_missions)) %>% 
  select(name)

top_astronauts <- unique(names)[1:10, "name"] %>% pull()

plot <- multiple_missions %>% 
  filter(name %in% top_astronauts, mission_number != 1) %>%
  mutate(mission_number = as.factor(mission_number)) %>% 
  ggplot(aes(fill = mission_number, values = on_earth)) +
  geom_waffle(color = "white", size = 2, flip = TRUE) +
  facet_wrap(~name, ncol = 1, strip.position = "left") +
  ggthemes::scale_fill_tableau(name="Mission Number") +
  coord_equal(ratio = 0.3) +
  labs(title = "What is the waiting time between mission for\n the top 10 astronauts with the most number of missions?",
       caption = "1 box = 1 year") +
  theme_minimal(base_family = "Roboto Condensed") +
  theme(plot.title = element_text(vjust = 0, hjust = 0),
        plot.caption = element_text(hjust = 1),
        strip.text.y.left = element_text(angle = 0, vjust = 0, hjust = 0),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE))

plot

ggsave("2020/plots/week_29.png", plot, device = "png", width = 10, height = 5)

```

## Appendix - Data Issues

1. For the following two astronauts:

- Franco Malerba
- Thomas, Andrew S. W.

The `year_of_mission` is before the `year_of_selection`

```{r}
astronauts_data %>% 
  filter(mission_number == 1) %>% 
  mutate(years_waiting = year_of_mission - year_of_selection) %>% 
  filter(years_waiting < 0)
```

2. For astronaut **Williams, Sunita L.** the `total_number_of_missions` seems incorrect. It should be 2 but instead it is 1 and 2.

3. For astronaut **Shepard, Alan B., Jr** the `total_number_of_missions` is given as 2 but there is only 1 entry.

