---
title: "#TidyTuesday 2002 week 33"
author: Kaustav Sen
date: "20 August 2022"
format: 
  html:
    code-overflow: wrap
    toc: true
    toc-depth: 2
    toc-location: left
---

## Loading packages

```{r}
#| message: false
library(tidyverse)  # for all the data wrangling 
library(magick)  # getting the images of the characters
library(gggrid)  # adding images and character names to the plots
```

## Reading in the data

```{r}
#| message: false
characters <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-08-16/characters.csv')
psych_stats <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-08-16/psych_stats.csv')
```

## Data wrangling

Will be working with the **Lord of the Rings** characters in this week's data. Initial idea is to compare top 10 traits across all the main characters provided in the data-set.

Here I have extracted the *top 10* traits based on those which show the greatest amount of variance amongst the characters. To compute the variance, first converted the `personality` column to a factor and then took its numeric representation.

```{r}
top_questions <-
  psych_stats |>
  filter(uni_name == "Lord of the Rings") |> 
  group_by(question) |> 
  mutate(
    question_var = as.factor(personality) |> as.numeric(personality) |> var()
  ) |> 
  ungroup() |> 
  filter(question_var == max(question_var)) |> 
  slice_head(n = 10) |> 
  pull(question)
```

```{r}
top_questions
```

Now that we have the traits by which we want to compare, lets filter the data-set and convert it to a plot-ready data frame.

But before we do that, we will use Gandalf's personality traits as the *anchor* points.

:::{.callout-note} 
Choosing Gandalf's personality traits as the anchor points was just an arbitrary decision. Could equally well have chosen any other character or based it on the `personality` column as well.
:::
It was tricky interpreting the `avg_rating` column, so thanks to [Tanya](https://twitter.com/tanya_shapiro) for the below explanation:

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">Tip for <a href="https://twitter.com/hashtag/TidyTuesday?src=hash&amp;ref_src=twsrc%5Etfw">#TidyTuesday</a> folks this week. Avg rating is tied to personality extremes. One idea might be to rescale it 🙂 <a href="https://t.co/bzuTcvtKtY">https://t.co/bzuTcvtKtY</a></p>&mdash; Tanya Shapiro (@tanya_shapiro) <a href="https://twitter.com/tanya_shapiro/status/1559506447070887939?ref_src=twsrc%5Etfw">August 16, 2022</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script> 

```{r}
anchor_traits <- 
  psych_stats |> 
  filter(char_name == "Gandalf", question %in% top_questions) |> 
  select(question, anchor = personality)

anchor_traits
```


```{r}
plot_df <- 
  psych_stats |>
  filter(uni_name == "Lord of the Rings", question %in% top_questions) |> 
  # Here I am retaining only those columns which I will be using in plotting.
  # This is not necessary - just helps viewing all relevant columns in one go 
  # when working interactively
  select(char_name, question, personality, avg_rating) |> 
  arrange(question) |>
  left_join(anchor_traits, by = "question") |> 
  mutate(rescaled_score = if_else(anchor == personality, avg_rating, 100 - avg_rating))
```

## Plotting

Now we have a data also prepped and ready. So, let's move onto some plotting!!

I was inspired by [Nicola's](https://twitter.com/nrennie35) and [Dan's](https://twitter.com/danoehm) submissions for this week to try and create mobile-friendly *card*-style plots. 

Definitely worth checking out their stunning plots as well:

:::{.grid}

:::{.g-col-5}
<blockquote class="twitter-tweet"><p lang="en" dir="ltr">This week&#39;s <a href="https://twitter.com/hashtag/TidyTuesday?src=hash&amp;ref_src=twsrc%5Etfw">#TidyTuesday</a> is about personality test results - I focused on Killing Eve characters! Inspired by <a href="https://twitter.com/bear_jordan_?ref_src=twsrc%5Etfw">@bear_jordan_</a> to try mobile-friendly data viz formats! Thanks to <a href="https://twitter.com/tanya_shapiro?ref_src=twsrc%5Etfw">@tanya_shapiro</a> for the data and plot inspiration!<br><br>Code: <a href="https://t.co/XiCurLLHFI">https://t.co/XiCurLLHFI</a><a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> <a href="https://twitter.com/hashtag/DataViz?src=hash&amp;ref_src=twsrc%5Etfw">#DataViz</a> <a href="https://t.co/ZQeoD6xcoE">pic.twitter.com/ZQeoD6xcoE</a></p>&mdash; Nicola Rennie (@nrennie35) <a href="https://twitter.com/nrennie35/status/1559513805440385027?ref_src=twsrc%5Etfw">August 16, 2022</a></blockquote>
:::

:::{.g-col-5}
<blockquote class="twitter-tweet"><p lang="en" dir="ltr"><a href="https://twitter.com/hashtag/TidyTuesday?src=hash&amp;ref_src=twsrc%5Etfw">#TidyTuesday</a> week 33: Psychometrics 🧠<br><br>Psychometric profiles for <a href="https://twitter.com/BreakingBad?ref_src=twsrc%5Etfw">@BreakingBad</a> main characters. Inspired by <a href="https://twitter.com/nrennie35?ref_src=twsrc%5Etfw">@nrennie35</a>&#39;s style I&#39;ve made individual character cards<br><br>🧵[1/3]<a href="https://twitter.com/hashtag/RStats?src=hash&amp;ref_src=twsrc%5Etfw">#RStats</a> <a href="https://twitter.com/hashtag/DataVisualization?src=hash&amp;ref_src=twsrc%5Etfw">#DataVisualization</a> <a href="https://twitter.com/hashtag/dataviz?src=hash&amp;ref_src=twsrc%5Etfw">#dataviz</a> <a href="https://t.co/5VXDpSdWZ6">pic.twitter.com/5VXDpSdWZ6</a></p>&mdash; Dan Oehm 🌲⛰️ (@danoehm) <a href="https://twitter.com/danoehm/status/1560405471085084674?ref_src=twsrc%5Etfw">August 18, 2022</a></blockquote>
:::

:::

For each character I wanted to create a bar-(*code*)-plot similar to how trading cards look like. I ended up creating two functions to re-use in creating the individual plots:

- `img_circle()` this function uses the `{grid}` package in order to add the character image and name at the top of the plot. This is used in conjunction with `grid_panel()` in the main plotting function `draw_barcode_plot()`

- `draw_barcode_plot` this is the main function which uses `{ggplot}` to create the plot

Expand the below section to view the full code.

```{r}
#| code-fold: true
#| code-overflow: scroll

# function to add image and character's name to the plot
img_circle <- function(data, coords) {
  img <- image_read(data$img_link[1])
  character_name <- data$name_label[1]

  outline <- circleGrob(x = 0.4, y = 1.1, r = 0.11, gp = gpar(col = "#F2A444", lwd = 3))
  image <- rasterGrob(img, vp = viewport(x = unit(0.4, "npc"),
                                         y = unit(1.1, "npc"),
                                         width = unit(0.2, "npc"),
                                         height = unit(0.2, "npc"),
                                         clip = circleGrob()))
  label <- textGrob(character_name, x = unit(0.53, "npc"), y = unit(1.1, "npc"),
                    hjust = 0, gp = gpar(fontfamily = "Cabinet Grotesk",
                                         fontface = "bold",
                                         fontsize = 13,
                                         col = "#F2A444"))
  gTree(children = gList(outline, image, label))
}

# main function to create the plot
draw_barcode_plot <- function(df, character_name) {
  plot_df_filtered <- 
  df |> 
  filter(char_name == character_name) |> 
  mutate(
    full = 100,
    anchor = fct_reorder(anchor, rescaled_score)
  ) |> 
  arrange(anchor)

  img_link <- characters |> 
    filter(name == character_name) |> 
    pull(image_link)

  num_labels <- str_c(
    round(plot_df_filtered$rescaled_score,0), 
    c(rep("", 9), "%")
  )
  
  subtitle <- "Using data from the Open-Source Psychometrics Project, we 
  analyse ten personality traits for each character from the Lord of the 
  Rings franchise" |> 
    str_wrap(65)
  
  ggplot(plot_df_filtered, aes(y = anchor)) +
    geom_col(aes(x = full), width = 0.4, fill = "grey90") +
    geom_col(aes(x = rescaled_score), width = 0.4, fill = "#A65F46") +
    geom_vline(xintercept = seq(0, 100, 2), color = "white") +
    annotate("text", x = 0, y = 1:10 + 0.38, 
             label = str_to_sentence(plot_df_filtered$anchor),
             hjust = 0, fontface = "bold", family = "Cabinet Grotesk", 
             size = 3.5, color = "grey20") +
    annotate("text", x = 100, y = 1:10 + 0.38, label = num_labels,
             hjust = 1, fontface = "bold", family = "Tabular", size = 3.5,
             color = "#A65F46") +
    grid_panel(img_circle, aes(img_link = img_link, 
                               name_label = character_name)) +
    labs(
      title = "Lord of the personalities",
      subtitle = subtitle
    ) +
    scale_x_continuous(limits = c(0, 100)) +
    coord_cartesian(clip = "off") +
    theme_minimal(base_size = 12, base_family = "Cabinet Grotesk") +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(face = "bold", size = rel(1.8), hjust = 0.5,
                                color = "#1B3C59"),
      plot.subtitle = element_text(size = rel(0.7), hjust = 0.5,
                                   margin = margin(b = 70), color = "grey45"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      axis.title = element_blank(),
      axis.text = element_blank()
    )
}
```

...and here are the final plots.

:::{.panel-tabset}

## Aragorn

```{r}
#| label: plot-aragorn
#| fig-width: 4
#| fig-asp: 1.62
#| dpi: 150

draw_barcode_plot(plot_df, "Aragorn")
```

## Gandalf

```{r}
#| label: plot-gandalf
#| fig-width: 4
#| fig-asp: 1.62
#| dpi: 150

draw_barcode_plot(plot_df, "Gandalf")
```

## Frodo

```{r}
#| label: plot-frodo
#| fig-width: 4
#| fig-asp: 1.62
#| dpi: 150

draw_barcode_plot(plot_df, "Frodo Baggins")
```

## Legolas

```{r}
#| label: plot-legolas
#| fig-width: 4
#| fig-asp: 1.62
#| dpi: 150

draw_barcode_plot(plot_df, "Legolas")
```

## Gimli

```{r}
#| label: plot-gimli
#| fig-width: 4
#| fig-asp: 1.62
#| dpi: 150

draw_barcode_plot(plot_df, "Gimli")
```

:::


